\SW: I. Crowd-sourcing the Discovery of Gravitational Lenses

### 2015-04-21 doc/sw-system.tex
Working through Anu's comments, up to Section 6 #112

At this point, the distribution of lens and source properties are quite
realistic. We then chose to make cuts of the simulated lens pool, in order to
reject sims that were likely to be diﬃcult to spot. While this ensured the
educational value of the sims when feedback was given, it meant that we should
only expect to detect real lens candidates that also pass these cuts. Any real
lens candidates we ﬁnd with properties outside the ranges of the sims, may in
some sense be considered bonuses, a product of the {\it imagination} of the
crowd in extrapolating their understanding of gravitational lenses into a regime
where they have not seen any before. In the \cfhtls project, the most important
selection cuts were to only keep sims with a)  second brightest image brighter
that $i = 23$ and b) total combined image magnitude brighter than 19--20
(\PaperTwo, Table~1). The combination of these two cuts results in a sample of
visible lensed image configurations.

@@

\subsubsection{\StageTwo: Refinement}

@@

We first quantify the performance of the \SW system in terms of the recovery of
the training set images. At \StageOne, this set contains around 5712 simulated
lenses, and 450 duds; at \StageTwo, we used 152 simulated lenses and 201 duds.
\Fref{fig:results:sample:roc} shows receiver operating characteristic (ROC)
curves for \cfhtls \StageOne and \StageTwo. These plots show the true positive
rate (TPR, the number of sims correctly detected divided by the total number of
sims in the training set), and the false positive rate (FPR, the number of duds
incorrectly detected  divided by the total number of duds in the training set), both
for a given sample of detections defined by a particular probability threshold,
which varies along the curves.  In both stages, these curves show that true
positive rates of  around 90\% were achieved, at very low false positive rates.

For comparison we show the results of an analysis where the classifications of
training images were ignored, and none of the agents allowed to learn: the
agents were instead simply assigned initial values of their confusion matrices
of 0.75 for each element, which then remained constant. This set-up emulates a
very simple unweighted voting scheme, where all classifications are treated
equally.  In this case, the TPR remains under 80\% in \StageOne and 60\% in
\StageTwo, thus quantifying  the benefit of including training images and
allowing the agents to learn via their Bayesian updates.

When the agents are allowed to update their confusion matrices, the choice of
initial confusion matrix is not very important. The same 0.75 initial values
applied to normal, learning agents resulted in only a slightly lower TPR than
the default case, nevertheless indicating that initializing the system less
conservatively results in slightly worse performance.

@@

At \StageTwo, where no retirement was carried out, it was possible to reach
100\% purity: the knee of the curve is at just under 90\% completeness. However,
the purity decreases rapidly if higher completeness than this is sought. The
optimal sample in this simulated lens search experiment would have been
constructed with a threshold value of $\pr(\LENS|C,T) > 0.47$. At 100\% purity
and 90\% completeness, it would have contained around 89 lens candidates.

@@

We are also grateful to Stuart Lynn, Kelly Borden, Laura Whyte, Brooke Simmons,
David Hogg, Thomas Jennings, Layne  Wright, Cecile Faure, Jonathan Coles, Stuart
Lowe and Jean-Paul Kneib for many useful conversations about citizen science and
gravitational lens detection, and to the Dark Energy Survey and Pan-STARRS strong
lensing science teams for their suggestions and encouragement.

PJM and ES thank the Institute of Astronomy and Astrophysics, Academia Sinica
(ASIAA) and Taiwan's Ministry of Science and Technology (MOST) for their
financial support of the workshop ``Citizen Science in Astronomy'' in March
2014, at which some parts of the SWAP analysis was developed.

>>>>

At this point, the distribution of lens and source properties are quite
realistic. We then chose to make cuts of the simulated lens pool, in order to
reject sims that were likely to be diﬃcult to spot. While this ensured the
educational value of the sims when feedback was given, it meant that we should
only expect to detect real lens candidates that also pass these cuts. Any real
lens candidates we ﬁnd with properties outside the ranges of the sims, may in
some sense be considered bonuses, a product of the {\it imagination} of the
crowd in extrapolating their understanding of gravitational lenses into a regime
where they have not seen any before. In the \cfhtls project, the most important
selection cuts were to only keep sims with a) second brightest image brighter
that $i = 23$ and b) total combined image magnitude fainter than 19--20
(\PaperTwo, Table~1). The combination of these two cuts results in a sample of
visible lensed image configurations.

@@

\subsubsection{\StageTwo: Refinement}
\label{sec:design:stages:two}

@@

We first quantify the performance of the \SW system in terms of the recovery of
the training set images. At \StageOne, this set contained around 5712 simulated
lenses, and 450 duds; at \StageTwo, we used 152 images of simulated lenses and
201 duds selected as \StageOne false positives (\Sref{sec:design:stages:two}).
\Fref{fig:results:sample:roc} shows receiver operating characteristic (ROC)
curves for \cfhtls \StageOne and \StageTwo. These plots show the true positive
rate (TPR, the number of sims correctly detected divided by the total number of
sims in the training set), and the false positive rate (FPR, the number of duds
incorrectly detected  divided by the total number of duds in the training set), both
for a given sample of detections defined by a particular probability threshold,
which varies along the curves.  In both stages, these curves show that true
positive rates of  around 90\% were achieved, at very low false positive rates.

For comparison we show the results of an analysis where the classifications of
training images were ignored, and the agents' confusion matrix elements
were instead all simply assigned initial values of 0.75
which then remained constant. This set-up emulates a
very simple unweighted voting scheme, where all classifications are treated
equally.  In this case, the TPR remains under 80\% in \StageOne and 60\% in
\StageTwo, thus illustrating the benefit of including training images and
allowing the agents to learn via their Bayesian updates.

When the agents are allowed to update their confusion matrices, the choice of
initial confusion matrix is not very important: the same 0.75 initial values
applied to normal, learning agents resulted in only a slightly lower TPR than
the default case.

@@

At \StageTwo, where no retirement was carried out, it was possible to reach
100\% purity: the knee of the curve is at just under 90\% completeness. However,
the purity decreases rapidly if higher completeness than this is sought. The
optimal sample in this simulated lens search experiment would have been
constructed with a threshold value of $\pr(\LENS|C,T) > 0.47$. At 100\% purity
and 90\% completeness, it would have contained around 89 lens candidates.
However, we remind the reader that all these values are dependent on the
properties of the training subjects, which were chosen to be fairly
visible: we expect the completeness to real lenses to be somewhat lower, as a
result.

@@

We thank Stuart Lynn, Kelly Borden, Laura Whyte, Brooke Simmons,
David Hogg, Thomas Jennings, Layne  Wright, Cecile Faure, Jonathan Coles, Stuart
Lowe and Jean-Paul Kneib for many useful conversations about citizen science and
gravitational lens detection, and to the Dark Energy Survey and Pan-STARRS strong
lensing science teams for their suggestions and encouragement.

PJM and ES also thank the Institute of Astronomy and Astrophysics, Academia Sinica
(ASIAA) and Taiwan's Ministry of Science and Technology (MOST) for their
financial support of the workshop ``Citizen Science in Astronomy'' in March
2014, at which some parts of the SWAP analysis was developed.
